---
title: "Priming and suppression of soil respiration from Ure2 incubation"
author: "Greg McKittrick"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load necessary library
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(emmeans)
```

## Set output directory value

```{r}
# Get current date
current_date <- Sys.Date()

# Format date as yyyy-mm-dd
date_string <- format(current_date, "%Y-%m-%d")

# Create output folder path
output_folder <- paste0("../Output/", date_string, "/")
```

## Read in Data

```{r}
# load in 13C data from GHG-IRMS
C13Data_filtered <- read.csv(file = paste0 (output_folder, "13CO2_partitioning.csv"))

# Convert variables to appropriate types
C13Data_filtered$Jar <- as.factor(C13Data_filtered$Jar)
C13Data_filtered$Rep <- as.factor(C13Data_filtered$Rep)
C13Data_filtered$Treatment <- as.factor(C13Data_filtered$Treatment)
C13Data_filtered$Case <- as.factor(C13Data_filtered$Case)
#C13Data_filtered$Day <- as.numeric(C13Data_filtered$Day)
```

## Calculate cumulative respiration attribted to substrates and soil

```{r}
C13Data_filtered <- C13Data_filtered %>% 
  arrange(Rep, Treatment , Day) %>%
  group_by(Rep, Treatment) %>%
  mutate(cumulative_respired_substrate_ugCO2_C_per_gdried_soil = cumsum(respired_substrate_ugCO2_C_per_gdried_soil),
         cumulative_soil_respired_ugCO2_C_per_gdried_soil = cumsum(soil_respired_ugCO2_C_per_gdried_soil)) %>% 
  arrange(Treatment)

```

## Plot cumulative respiration attributed to soil (no averaging of reps)

```{r}
#Plot
p<- ggplot(C13Data_filtered, aes(x = Day, y = cumulative_soil_respired_ugCO2_C_per_gdried_soil, color = Treatment)) +
  geom_jitter(alpha = 0.5, width = 0.2) +
  labs(x = "Days", y = "ug CO2 C per g dried soil", title = "Cumulative soil derived respiration") +
  theme_minimal()
  #facet_grid(Treatment~Rep)+
  #geom_hline(yintercept = 341.63003, color = "red", linetype = "dashed", size = 0.2)
p

ggsave(file = paste0(output_folder,"Cumulative_Soil_Derived_Respiration.pdf"), plot = p)
```

## Create a summary of cumulative_soil_respired_ugCO2_C 

```{r}
summaryC13Data_filtered <- C13Data_filtered %>%
  group_by(Treatment, Day) %>%
  summarise(
    avg_cumulative_soil_respired_ugCO2_C_per_gdried_soil = mean(cumulative_soil_respired_ugCO2_C_per_gdried_soil),
    n = n(),
    st_dev = sd(cumulative_soil_respired_ugCO2_C_per_gdried_soil),
    .groups = 'drop'
  ) %>%
  group_by(Treatment) %>%
  mutate(cumulative_stdDev = sqrt(cumsum(st_dev ^ 2))) %>%
  arrange(Treatment)
```

## Plot cumulative respiration attributed to soil (Reps averaged)

```{r}
p <- ggplot(summaryC13Data_filtered, aes(x = Day, y = avg_cumulative_soil_respired_ugCO2_C_per_gdried_soil, color = Treatment)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(ymin = avg_cumulative_soil_respired_ugCO2_C_per_gdried_soil - cumulative_stdDev, ymax = avg_cumulative_soil_respired_ugCO2_C_per_gdried_soil + cumulative_stdDev), width = 0.2) +
  geom_segment(aes(x = Day - 0.4, xend = Day + 0.4, y = avg_cumulative_soil_respired_ugCO2_C_per_gdried_soil + cumulative_stdDev, yend = avg_cumulative_soil_respired_ugCO2_C_per_gdried_soil + cumulative_stdDev, color = Treatment)) +
  geom_segment(aes(x = Day - 0.4, xend = Day + 0.4, y = avg_cumulative_soil_respired_ugCO2_C_per_gdried_soil - cumulative_stdDev, yend = avg_cumulative_soil_respired_ugCO2_C_per_gdried_soil - cumulative_stdDev, color = Treatment)) +
  labs(x = "Days", y = "ugCO2 C per g dried soil\n+/- cumulative sd", title = " Average cumulative respiration from soil") +
  theme_minimal()
  #facet_grid(Treatment~Rep)+
  #geom_hline(yintercept = 341.63003, color = "red", linetype = "dashed", size = 0.2)

p

ggsave(file = paste0(output_folder,"Summary_Cumulative_Soil_Derived_Respiration.pdf"), plot = p)

```

## Plot cumulative respiration attributed to substrate (no averaging of reps)

```{r}
#Plot
p<- ggplot(C13Data_filtered, aes(x = Day, y = cumulative_respired_substrate_ugCO2_C_per_gdried_soil, color = Treatment)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(x = "Days", y = "ugCO2 C per g dried soil", title = "Cumulative Respiration\nfrom Substrate") +
  theme_minimal()
p

ggsave(file = paste0(output_folder,"Cumulative_Substrate_Derived_Respiration.pdf"), plot = p)
```

## Create a summary of substrate_cumulative_repsiration

```{r}
summaryC13Data_filtered2 <- C13Data_filtered %>% 
  group_by(Treatment, Day) %>% 
  summarise(avg_cumulative_respired_substrate_ugCO2_C_per_gdried_soil = mean(cumulative_respired_substrate_ugCO2_C_per_gdried_soil),
            n = n(),
            st_dev = sd(cumulative_respired_substrate_ugCO2_C_per_gdried_soil),
    .groups = 'drop'
  ) %>%
  group_by(Treatment) %>%
  mutate(cumulative_stdDev = sqrt(cumsum(st_dev ^ 2))) %>%
  arrange(Treatment)
```

## Plot summary cumulative respiration attributed to substrate (reps averaged)

```{r}
p <- ggplot(summaryC13Data_filtered2, aes(x = Day, y = avg_cumulative_respired_substrate_ugCO2_C_per_gdried_soil, color = Treatment)) +
  geom_point() +
  geom_errorbar(aes(ymin = avg_cumulative_respired_substrate_ugCO2_C_per_gdried_soil - cumulative_stdDev, ymax = avg_cumulative_respired_substrate_ugCO2_C_per_gdried_soil + cumulative_stdDev), width = 0.2) +
  geom_segment(aes(x = Day - 0.5, xend = Day + 0.5, y = avg_cumulative_respired_substrate_ugCO2_C_per_gdried_soil - cumulative_stdDev, yend = avg_cumulative_respired_substrate_ugCO2_C_per_gdried_soil - cumulative_stdDev, color = Treatment)) +
  geom_segment(aes(x = Day - 0.5, xend = Day + 0.5, y = avg_cumulative_respired_substrate_ugCO2_C_per_gdried_soil + cumulative_stdDev, yend = avg_cumulative_respired_substrate_ugCO2_C_per_gdried_soil + cumulative_stdDev, color = Treatment)) +
  labs(x = "Days", y = "ugCO2 C per gdried soil\n +/- cumulative sd", title = "Average cumulative\nrespiration from substrate") +
  theme_minimal()

p

ggsave(file = paste0(output_folder,"Summary_Cumulative_Substrate_Derived_Respiration.pdf"), plot = p)
```


## Calculate daily soil respiration averages

```{r}

daily_soil_only_respired <- C13Data_filtered %>% 
  filter(Treatment == "Soil Only") %>% 
  group_by(Day) %>% 
  summarise(n = n(),
            daily_soil_only_respired_ug_CO2_C_per_gdried_soil = mean(soil_respired_ugCO2_C_per_gdried_soil),
            st_dev = sd(soil_respired_ugCO2_C_per_gdried_soil))

all_soil_only_respired <- C13Data_filtered %>% 
  filter(Treatment == "Soil Only") %>% 
  summarise(all_soil_only_respired_ug_CO2_C_per_gdried_soil = mean(soil_respired_ugCO2_C_per_gdried_soil))
```

# Merge daily_soil_only_respired into merged_13C_licor_substrates_soil. 

```{r}
# Merge daily_soil_only_respired into merged_13C_licor_substrates_soil
C13Data_filtered <- C13Data_filtered %>%
  left_join(daily_soil_only_respired, by = "Day") %>% 
  select(Jar, Case, Day, Rep, Treatment, CO2_delta_13C_VPDB, substrate_C_added_ug, delta_C13_substrates, dried_soil_g, ugCO2_C_per_gdried_soil, cumulative_ugCO2_C_per_gdried_soil, daily_soil_only_mean_CO2_delta_13C_VPDB, proportion_substrate_respiration, final_proportion_substrate_respiration, respired_substrate_ugCO2_C_per_gdried_soil, soil_respired_ugCO2_C_per_gdried_soil, ug_C_per_g_dried_soil, percent_substrate_respired, cumulative_percent_substrate_respired, cumulative_respired_substrate_ugCO2_C_per_gdried_soil, cumulative_soil_respired_ugCO2_C_per_gdried_soil, daily_soil_only_respired_ug_CO2_C_per_gdried_soil)

```

## Calculate % of daily soil respiration for all jars

```{r}
C13Data_filtered <- C13Data_filtered %>% 
  mutate(percent_of_daily_soil_only_mean = (soil_respired_ugCO2_C_per_gdried_soil / daily_soil_only_respired_ug_CO2_C_per_gdried_soil) *100)

```

## Plot % of daily soil respiration (no averaging of reps)

```{r}
p <- ggplot(C13Data_filtered,aes(x = Day, y = percent_of_daily_soil_only_mean, color = Treatment, shape = Rep )) +
  geom_line(stat = "identity") +
  #geom_errorbar(aes(ymin = cumulative_ugCO2_C_gdried_soil - cumulative_stdDev, ymax = cumulative_ugCO2_C_gdried_soil + cumulative_stdDev), width = 0.2) +
  labs(x = "Days", y = "Percent", title = "Percent of daily soil respiration ") +
  theme_minimal()+
  facet_grid(Treatment~Rep)+
  geom_hline(yintercept = 100, color = "red", linetype = "dashed", size = 0.2)+
  theme(
    plot.title = element_text(size = 20),     # Title font size
    axis.title.x = element_text(size = 15),   # X axis label font size
    axis.title.y = element_text(size = 15),   # Y axis label font size
    strip.text.y = element_blank()            # Remove facet label for Treatment)
  )


p

ggsave(file = paste0(output_folder,"Percent_of_daily_soil_respiration.pdf"), plot = p)
```

## Create summary of % of daily soil cumulative respiration

```{r}
summaryC13Data_filtered3 <- C13Data_filtered %>% 
  group_by(Treatment, Day) %>% 
  summarise(avg_percent_of_daily_soil_only_mean = mean(percent_of_daily_soil_only_mean),
            n = n(),
            st_dev = sd(percent_of_daily_soil_only_mean))
```

## Plot summary  % of daily soil respiration (reps averaged)

```{r}
p <- ggplot(summaryC13Data_filtered3,aes(x = Day, y = avg_percent_of_daily_soil_only_mean, color = Treatment)) +
  geom_point(stat = "identity", size = 0.5) +
  geom_errorbar(aes(ymin = avg_percent_of_daily_soil_only_mean - st_dev, ymax = avg_percent_of_daily_soil_only_mean + st_dev), width = 0.2) +
  labs(x = "Days", y = "Percent\n+/- sd", title = " Average % of daily soil respiration ") +
  theme_minimal()+
  facet_wrap(~Treatment)+
  geom_hline(yintercept = 100, color = "red", linetype = "dashed", size = 0.2)+
  theme(
    plot.title = element_text(size = 20),     # Title font size
    axis.title.x = element_text(size = 15),   # X axis label font size
    axis.title.y = element_text(size = 15),   # Y axis label font size
    strip.text.y = element_blank()            # Remove facet label for Treatment)
  )


p

ggsave(file = paste0(output_folder,"Summary_Percent_of_daily_soil_respiration.pdf"), plot = p)

```

## Plot % of daily soil respiration (reps averaged)

```{r}
p <- ggplot(summaryC13Data_filtered3, aes(x = Day, y = avg_percent_of_daily_soil_only_mean, color = Treatment)) +
  geom_line(stat = "identity") +
  geom_errorbar(aes(ymin = avg_percent_of_daily_soil_only_mean - st_dev, ymax = avg_percent_of_daily_soil_only_mean + st_dev), width = 0.1) +
  geom_segment(aes(x = Day - 0.2, xend = Day + 0.2, y = avg_percent_of_daily_soil_only_mean + st_dev, yend = avg_percent_of_daily_soil_only_mean + st_dev, color = Treatment)) +
  geom_segment(aes(x = Day - 0.2, xend = Day + 0.2, y = avg_percent_of_daily_soil_only_mean - st_dev, yend = avg_percent_of_daily_soil_only_mean - st_dev, color = Treatment)) +
  labs(x = "Days", y = "Percent", title = "Average % of daily soil respiration") +
  theme_minimal()

p
```
## Extract cumulative daily soil respiration averages

```{r}

daily_soil_only_cumulative <- C13Data_filtered %>% 
  filter(Treatment == "Soil Only") %>% 
  group_by(Day) %>% 
  summarise(n = n(),
            daily_cumulative_soil_only_respired_ug_CO2_C_per_gdried_soil = mean(cumulative_soil_respired_ugCO2_C_per_gdried_soil),
            st_dev = sd(cumulative_soil_respired_ugCO2_C_per_gdried_soil))

```

# Merge cum_daily_soil_only_respired into cumulative_respiration

```{r}
# Merge daily_soil_only_respired into merged_13C_licor_substrates_soil
C13Data_filtered <- C13Data_filtered %>%
  left_join(daily_soil_only_cumulative, by = "Day") %>% 
  select(Jar, Case, Day, Rep, Treatment, CO2_delta_13C_VPDB, substrate_C_added_ug, delta_C13_substrates, dried_soil_g, ugCO2_C_per_gdried_soil, cumulative_ugCO2_C_per_gdried_soil, daily_soil_only_mean_CO2_delta_13C_VPDB, proportion_substrate_respiration, final_proportion_substrate_respiration, respired_substrate_ugCO2_C_per_gdried_soil, soil_respired_ugCO2_C_per_gdried_soil, ug_C_per_g_dried_soil, percent_substrate_respired, cumulative_percent_substrate_respired, cumulative_respired_substrate_ugCO2_C_per_gdried_soil, cumulative_soil_respired_ugCO2_C_per_gdried_soil, daily_soil_only_respired_ug_CO2_C_per_gdried_soil, percent_of_daily_soil_only_mean, daily_cumulative_soil_only_respired_ug_CO2_C_per_gdried_soil)
```

## Calculate % of daily soil only cumulative respiration for all jars

```{r}
C13Data_filtered <- C13Data_filtered %>% 
  mutate(percent_of_daily_cumulative_soil_only = (cumulative_soil_respired_ugCO2_C_per_gdried_soil / daily_cumulative_soil_only_respired_ug_CO2_C_per_gdried_soil) *100)

```

## Plot % of daily soil respiration (no averaging of reps)

```{r}
p <- ggplot(C13Data_filtered,aes(x = Day, y = percent_of_daily_cumulative_soil_only, color = Treatment, shape = Rep )) +
  geom_line(stat = "identity") +
  #geom_errorbar(aes(ymin = cumulative_ugCO2_C_gdried_soil - cumulative_stdDev, ymax = cumulative_ugCO2_C_gdried_soil + cumulative_stdDev), width = 0.2) +
  labs(x = "Days", y = "Percent", title = "Percent of daily\ncumulative soil respiration ") +
  theme_minimal()+
  facet_grid(Treatment~Rep)+
  geom_hline(yintercept = 100, color = "red", linetype = "dashed", size = 0.2)

p

ggsave(file = paste0(output_folder,"Percent_of_daily_cumulative_soil_respiration.pdf"), plot = p)

```

## Plot % of daily soil cumulative respiration (no averaging of reps)

```{r}
p <- ggplot(C13Data_filtered,aes(x = Day, y = percent_of_daily_cumulative_soil_only, color = Treatment, shape = Rep )) +
  geom_point(stat = "identity") +
  #geom_errorbar(aes(ymin = cumulative_ugCO2_C_gdried_soil - cumulative_stdDev, ymax = cumulative_ugCO2_C_gdried_soil + cumulative_stdDev), width = 0.2) +
  labs(x = "Days", y = "Perent", title = "Percent of daily soil cumulative respiration ") +
  theme_minimal()+
  geom_hline(yintercept = 100, color = "red", linetype = "dashed", size = 0.2)

p
```

## Create summary of % of daily soil cumulative respiration

```{r}
summaryC13Data_filtered4 <- C13Data_filtered %>%  
  group_by(Treatment, Day) %>% 
  summarise(avg_percent_of_daily_cumulative_soil_only = mean(percent_of_daily_cumulative_soil_only),
            n = n(),
            st_dev = sd(percent_of_daily_cumulative_soil_only))
```

## Plot % of daily soil cumulative respiration (reps averaged)

```{r}
p <- ggplot(summaryC13Data_filtered4,aes(x = Day, y = avg_percent_of_daily_cumulative_soil_only, color = Treatment)) +
  geom_point(stat = "identity", size = 0.5) +
  geom_errorbar(aes(ymin = avg_percent_of_daily_cumulative_soil_only - st_dev, ymax = avg_percent_of_daily_cumulative_soil_only + st_dev), width = 0.2) +
  labs(x = "Days", y = "Percent\n+/- sd", title = " Average Percent of Daily Cumulative Soil Respiration ") +
  theme_minimal()+
  facet_wrap(~Treatment)+
  geom_hline(yintercept = 100, color = "red", linetype = "dashed", size = 0.2)+
  theme(
    plot.title = element_text(size = 20),     # Title font size
    axis.title.x = element_text(size = 15),   # X axis label font size
    axis.title.y = element_text(size = 15),   # Y axis label font size
    strip.text.y = element_blank()            # Remove facet label for Treatment)
  )

p

ggsave(file = paste0(output_folder,"Summary_Percent_of_daily_cumulative_soil_respiration.pdf"), plot = p)


```

## Plot % of daily soil cumulative respiration (reps averaged)

```{r}
p <- ggplot(summaryC13Data_filtered4, aes(x = Day, y = avg_percent_of_daily_cumulative_soil_only, color = Treatment)) +
  geom_line(stat = "identity") +
  geom_errorbar(aes(ymin = avg_percent_of_daily_cumulative_soil_only - st_dev, ymax = avg_percent_of_daily_cumulative_soil_only + st_dev), width = 0.1) +
  geom_segment(aes(x = Day - 0.2, xend = Day + 0.2, y = avg_percent_of_daily_cumulative_soil_only + st_dev, yend = avg_percent_of_daily_cumulative_soil_only + st_dev, color = Treatment)) +
  geom_segment(aes(x = Day - 0.2, xend = Day + 0.2, y = avg_percent_of_daily_cumulative_soil_only - st_dev, yend = avg_percent_of_daily_cumulative_soil_only - st_dev, color = Treatment)) +
  labs(x = "Days", y = "Percent", title = "Average % of daily soil cumulative respiration") +
  theme_minimal()

p
```