---
title: "Cumulative_CO2_C"
author: "Greg McKittrick"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load packages
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(emmeans)
```

## Create Output directory

```{r}

# Get current date
current_date <- Sys.Date()

# Format date as yyyy-mm-dd
date_string <- format(current_date, "%Y-%m-%d")

# Create output folder path
output_folder <- paste0("../Output/", date_string, "/")

# Create output folder if it doesn't exist
if (!file.exists(output_folder)) {
  dir.create(output_folder, recursive = TRUE)
}

# Now you can use the output_folder variable to save your output
# For example:
# saveRDS(your_output_data, file = paste0(output_folder, "output_data.rds"))
```



## load in data
```{r}
allData <- read_excel("../LICOR/20240321_Li-Cor-calculation.xlsx", sheet = 4)

#change column name and remove NA nows from day 25
allData <- allData %>%
  rename(ugCO2_C_per_gdried_soil = "ugCO2-C/gdried soil") %>% 
  na.omit()

# Convert variables to appropriate types
allData$Jar <- as.factor(allData$Jar)
allData$Rep <- as.factor(allData$Rep)
#allData$Day <- as.factor(allData$Day)
allData$Treatment <- as.factor(allData$Treatment)
allData$Case <- paste(allData$Treatment, allData$Rep, sep = "_")
allData$Case <- as.factor(allData$Case)
str(allData)
```

## Filter out bad observations due to techinical issues (leaky lids)


```{r}

# Step 1: Filter out specific Case for days 14, and 17, 22, 30, and 37
#once one has been removed all subseqeunt days need to be removed because cumulative_ugCO2_C_gdried_soil will be falsley low at those points without consideration of the missing one befor eit
allData_filtered <- allData %>%
  filter(!(Day %in% c("14", "17", "22","30", "37") & Case == "BSA_1"))

# Step 1: Filter out specific Case for days 30 and 37
allData_filtered <- allData_filtered %>%
  filter(!(Day %in% c("30","37") & Case == "Ure2_2"))
```


##Calculate cumulative ugCO2_C_per_gdried_soil

```{r}
cumulativeData <- allData_filtered %>%
  arrange(Rep, Treatment , Day) %>%
  group_by(Rep, Treatment) %>%
  mutate(cumulative_ugCO2_C_per_gdried_soil = cumsum(ugCO2_C_per_gdried_soil)) %>% 
  arrange(Treatment)

# Create a summary of the cumulative CO2 C data 
cumulativeDataSummary <- allData_filtered %>%
  arrange(Rep, Treatment , Day) %>%
  group_by(Rep, Treatment) %>%
  mutate(cumulative_ugCO2_C_per_gdried_soil = cumsum(ugCO2_C_per_gdried_soil)) %>% 
  group_by(Day, Treatment) %>% 
  summarise(n = n(),
    avg_cum_ugCO2_C_per_gdried_soil = mean(cumulative_ugCO2_C_per_gdried_soil),
            stdDev = sd(cumulative_ugCO2_C_per_gdried_soil),
            .groups = 'drop') %>% 
  group_by(Treatment) %>%
  mutate(cumulative_stdDev = sqrt(cumsum(stdDev^2))) %>% 
  arrange(Treatment)

#Save a csv file with this cumulative CO2 C 
write.csv(cumulativeDataSummary, file = paste0(output_folder, "summary_cumulativeData_CO2_C.csv"))

cumulativeDataForModel <- cumulativeData %>%
  mutate(Day = as.factor(Day))
  
  
```


##Run a lmer model with cumulative_ugCO2_C_gdried_soil

In summary, this model is investigating how the cumulative amount of CO2 per gram of dried soil (cumulative_ugCO2_C_per_gdried_soil) is influenced by Treatment, Day, and their interaction, while accounting for the variability between different cases (Case) by including random intercepts.

```{r}
# Be sure to define time as.factor!!!
#Case (takes 16 unique values corresponding to 16 jars!)

Model <- lmer(cumulative_ugCO2_C_per_gdried_soil ~ Treatment*Day + (1|Case), data = cumulativeDataForModel)
#Model <- lm(cumulative_ugCO2_C_gdried_soil ~ Treatment*Day, data = cumulativeDataForModel)
#Model <- aov(cumulative_ugCO2_C_gdried_soil ~ Treatment*Day, data = cumulativeDataForModel)

#Treatment and Day are fixed effects
#(1|Case) specifies case (or jar) as a random effect to account for repeated measures on the same jar over time.
#ANOVA table to test for main effect and interaction.  This is a common starting point. 
anova(Model)

# Run anova(Model) and capture the output
anova_output <- capture.output(anova(Model))

# Write the captured output to a text file
writeLines(anova_output, con = paste0(output_folder, "cumulative_CO2_C_aov_anova.txt"))


# Diagnostic plots to check assumptions of normality and equal variance.
plot(Model)


#emmeans is VERY flexible!
#emmeans_results<-emmeans(Model, pairwise ~ Treatment|Day)
emmeans_results <- emmeans(Model, pairwise ~ Treatment|Day)
emmeans_results
#The code above will compare the means for the treatments, at EACH time point.
# Capture the output of emmeans()
emmeans_output <- capture.output(summary(emmeans_results))

# Write the captured output to a text file
writeLines(emmeans_output, con = paste0(output_folder, "cumulative_CO2_C_aov_emmeans.txt"))

```

#Plot the summary data

```{r}
#Plot with error bars cumulative_stdDev
p<- ggplot(cumulativeDataSummary, aes(x = Day, y = avg_cum_ugCO2_C_per_gdried_soil, color = Treatment)) +
  geom_point() +
  geom_errorbar(aes(ymin = avg_cum_ugCO2_C_per_gdried_soil - cumulative_stdDev, ymax = avg_cum_ugCO2_C_per_gdried_soil + cumulative_stdDev), width = 0.2) +
  labs(x = "Days", y = "ug CO2 C per g soil", title = "Cumulative Average ug CO2 C\nper g soil") +
  theme_minimal()

p

# Save plot as PDF
ggsave( file = paste0(output_folder,"Cumulative_CO2_C.pdf"), plot = p)

# Extract contrast information into a data frame
contrast_df <- as.data.frame(emmeans_results$contrasts)

# Filter contrast_df to include only the desired contrasts
filtered_contrasts <- contrast_df %>% 
  filter(contrast == c("BSA - Ure2","Soil Only - Ure2") & Day == "37")

# Define a function to convert p-values to stars
p_to_stars <- function(p) {
  if (p < 0.001) {
    return("***")
  } else if (p < 0.01) {
    return("**")
  } else if (p < 0.05) {
    return("*")
  } else {
    return("")
  }
}

# Convert p-values to stars
filtered_contrasts$stars <- sapply(filtered_contrasts$p.value, p_to_stars)

#Plot with error bars cumulative_stdDev AND p values for "BSA - Ure2","Soil Only - Ure2" contrasts on Day 30
p <- ggplot(cumulativeDataSummary, aes(x = Day, y = avg_cum_ugCO2_C_per_gdried_soil, color = Treatment)) +
  geom_point() +
  geom_errorbar(aes(ymin = avg_cum_ugCO2_C_per_gdried_soil - cumulative_stdDev, ymax = avg_cum_ugCO2_C_per_gdried_soil + cumulative_stdDev), width = 0.2) +
  labs(x = "Days", y = "ug CO2 C per g soil", title = "Cumulative Average ug CO2 C\nper g soil") +
  theme_minimal()

# Add the first vertical line
p <- p +
  geom_segment(aes(x = 29, y = 336.27666, xend = 29, yend = 565.35026), color = "blue") +
  geom_text(aes(x = 27.5, y = 425, label = "***"), color = "blue", size = 3)

# Add the second vertical line
p <- p + 
  geom_segment(aes(x = 31, y = 284.94867, xend = 31, yend = 336.27666), color = "red") +
  geom_text(aes(x = 32.5, y = 270, label = "***
"), color = "red", size = 3)
p

```

##Read in substrate addition weights

```{r}
substrate_weights <- read_excel("../SetupDetails.xlsx", sheet = 1)

#change jar to factor and and change name of column
substrate_weights$Jar <- as.factor(substrate_weights$Jar)
substrate_weights <- substrate_weights %>%
  rename(substrate_C_added_ug = `amount C-added (ug)`)
```

##Add substrates addition data to cumulativeData

```{r}
cumulativeData <- cumulativeData %>%
  left_join(substrate_weights, by = "Jar", suffix = c("", "")) %>%
  select(Jar, Case, Day, Rep, Treatment, ugCO2_C_per_gdried_soil, cumulative_ugCO2_C_per_gdried_soil, substrate_C_added_ug)

```

##Calculate daily soil only average cumulative_ugCO2_C_gdried_soil

```{r}
daily_soil_only <- cumulativeData %>% 
  filter(Treatment == "Soil Only") %>% 
  group_by(Day) %>% 
  summarise(daily_soil_only_mean_ugCO2_C_per_gdried_soil = mean(cumulative_ugCO2_C_per_gdried_soil))

# Merge daily_soil_only_mean into cumulativeData_substrates
cumulativeData <- cumulativeData %>%
  left_join(daily_soil_only, by = "Day")
```

## Calculate ugCO2_C_per_gdried_soil - daily_soil_only_mean ,ug_CO2_C_per_ug_C_added, cumulative ug_CO2_C_per_ug_C_added

```{r}
cumulativeData <- cumulativeData %>%
  mutate(treatment_minus_daily_soil_only_mean_ugCO2_C_per_gdried_soil = cumulative_ugCO2_C_per_gdried_soil -daily_soil_only_mean_ugCO2_C_per_gdried_soil,
         ug_CO2_C_per_ug_C_added = treatment_minus_daily_soil_only_mean_ugCO2_C_per_gdried_soil / substrate_C_added_ug) %>% 
  group_by(Rep, Treatment) %>%
  mutate(cumulative_ug_CO2_C_per_ug_C_added = cumsum(ug_CO2_C_per_ug_C_added))

#Create a CSV file with all data and transformations
write.csv(cumulativeData, file = paste0(output_folder, "cumulativeData_CO2_C.csv"))

#create summary data frame with replicates averaged
cumulativeDataSummary2 <- cumulativeData %>% 
  group_by(Treatment, Day) %>% 
  summarise(n=n(),
            avg_cumulative_ug_CO2_C_per_ug_C_added = mean(cumulative_ug_CO2_C_per_ug_C_added),
            stdDev = sd(cumulative_ug_CO2_C_per_ug_C_added),
            .groups = 'drop') %>% 
  group_by(Treatment) %>%
  mutate(cumulative_stdDev = sqrt(cumsum(stdDev^2))) %>% 
  arrange(Treatment)

#Create a csv file with this summary of cumulative CO2 C per C_added
write.csv(cumulativeDataSummary2, file = paste0 (output_folder, "summary_cumulativeData_ug_CO2_C_per_ug_C_added.csv"))

```

##Plot cumulative ug_CO2_C_per_ug_C_added no averaging of reps

```{r}
#Plot
p<- ggplot(cumulativeData, aes(x = Day, y = cumulative_ug_CO2_C_per_ug_C_added, color = Treatment)) +
  geom_point() +
  #geom_errorbar(aes(ymin = avg_cum_ugCO2_C_gdried_soil - cumulative_stdDev, ymax = avg_cum_ugCO2_C_gdried_soil + cumulative_stdDev), width = 0.2) +
  labs(x = "Days", y = "ug CO2 C per ug C added", title = "Cumulative ug CO2 C per ug C added") +
  theme_minimal()


p
```

##Plot cumulative ug_CO2_C_per_ug_C_added reps averaged

```{r}

#Plot with error bars cumulative_stdDev
p<- ggplot(cumulativeDataSummary2, aes(x = Day, y = avg_cumulative_ug_CO2_C_per_ug_C_added, color = Treatment)) +
  geom_point() +
  geom_errorbar(aes(ymin = avg_cumulative_ug_CO2_C_per_ug_C_added - cumulative_stdDev, ymax = avg_cumulative_ug_CO2_C_per_ug_C_added + cumulative_stdDev), width = 0.2) +
  labs(x = "Days", y = "ug CO2 C per ug C added", title = "Average Cumulative ug\nCO2 C per ug C added") +
  theme_minimal()


p

#save the plot
ggsave( file = paste0 (output_folder, "cumulative_ug_CO2_C_per_ug_C_added.pdf"), p)
```



